Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2020-12-29 GWLDCII.tex
refs in past tense

Wilson loop diagrams, as we will study them, are objects given by a perturbative expansion of the holomorphic Wilson loop in twistor space.
Classically, the interactions of the theory were studied as maximal helicity violating diagrams ($MHV$ diagrams), next-to-maximal helicity violating diagrams and so on ($N^kMHV$ diagrams) in momentum space. Adamo and Mason, in \cite{Adamo:2011pr}, give an elegant formulation for the $N^kMHV$ diagrams in twistor space. The Wilson loop diagram perspective of this paper is dual to this twistor representation of $MHV$ diagrams.
The objects are closely related and carry the same information.  We choose to emphasize the dual perspective by our focus on the Wilson loop diagrams.

@@

In \cite{wilsonloop}, Agarwala and Marin-Amat uncovered a connection between the Wilson loop diagrams and positroids, defined as matroids that can be realized as an element of $\Gr(k,n)$. Other work has shown a connection between the volume filled by Wilson loop diagrams and the Amplituhedron \cite{casestudy, Amplituhedronsquared}. Specifically, Marcott shows that the space parametrized by the Wilson loop diagrams is  exactly the correct positroid cell in $\Gr(k,n)$, up to a set of measure zero \cite{WLDdim}.

In the above work and elsewhere, it is clear that while the space parametrized by the Wilson loop diagrams is related to the Amplituhedron, they are not the same space, nor is a direct connection between the geometry of the Amplituhedron and the geometry of the Wilson loop diagrams known. For instance, it is conjectured that the Amplituhedron is orientable, whereas Agarwala and Marcott show in \cite{non-orientable} that the space parametrized by the Wilson loop diagrams as a subspace of $\Gr(k, n+1)$ can be seen as a non-orientable vector bundle over some submanifold of $\Gr(k,n)$. Furthermore, the cells associated to the Wilson loop diagrams are different from the cells associated to BCFW diagrams, which can be seen by comparing the results of \cite{AmplituhedronDecomposition} and \cite{casestudy}. Furthermore, the spaces in the two situations have different dimensions, as we will discuss further below, and examples show \cite{HeslopStewart} that they do not relate by a coordinate projection or boundary map or other straightforward function.  None the less, from their common physical origin, it is believed that once the integrals over these spaces are taken the results will parametrize related spaces \cite{wilsonloop, Amplituhedronsquared}.

@@

This paper continues the program of establishing concrete connections between the combinatorics of the Wilson loop diagram as determined by the physics, and the combinatorics of the associated positroid cell as identified in \cite{wilsonloop}. In this paper, we give an algorithm to read the Grassmann necklace, one of the standard combinatorial characterizations of the positroid cell, directly off the Wilson loop diagram. We also show that the positroid cells associated to Wilson loop diagrams are all $3k$-dimensional, where $k$ is the number of propagators in the diagram; by contrast, the BCFW cells are all $4k$-dimensional \cite{Arkani-Hamed:2013jha}. In \cite{non-orientable} the authors note that the Wilson loop diagrams only correspond to positroid cells when the gauge vector is ignored. Including the gauge vector defines a $4k$ dimensional subspace of $\mathbb{G}_\R(k,n+1)$.

@@

The matrix $C(W)$ defines a positroid cell associated to $W$ in the positive Grassmannian $\Gr(n, |\cP|)$, which we denote by $\Sigma(W)$ \cite[Theorem 3.41]{wilsonloop}. Marcott \cite[Theorem 8.4]{WLDdim} shows that  the matrix $C(W)$ parametrizes a space whose closure is exactly the closure of $\Sigma(W)$. That is, the space defined by $C(W)$ and the cell $\Sigma(W)$ agree up to a set of measure $0$. Most of the work in this paper and in the previous paper in the series \cite{generalcombinatoricsI} focuses on characterizing the positroid cell associated to each Wilson loop diagram $W$.

@@

Postnikov shows in \cite{Postnikov} that the positroids of $\Gr(k,n)$ are indexed by many different collections of objects, each with their own advantages and disadvantages. The two most suited to our purposes are Grassmann necklaces \cite[section 16]{Postnikov} and Le diagrams \cite[section 6]{Postnikov}, which we discuss. In order to do so, we first need some preliminary definitions.

@@

In \cite{wilsonloop}, Agarwala and Marin-Amat show that $M(W)$ can also be read directly from the Wilson loop diagram itself.

@@

By relating the behavior of the propagators in $W$ to the matroidal properties of $M(W)$, Agarwala and Marin-Amat also showed that $M(W)$ is a positroid whenever $W$ is admissible \cite[Corollary 3.39]{wilsonloop}. \todo{Decision -- make them past ------ check tenses throughout (M.24) -- maybe past is better, see the ``had'' at the beginning of the next paragraph, can we discuss?}

@@

The flavour of the calculations is similar to \cite{CGdyck}\footnote{Thanks to a referee for pointing this paper out.} as in both cases the bijections are very explicit with a similar iterative flavour using Le diagrams and Grassmann necklaces related to another class of objects, for us Wilson loop diagrams and for them unit interval orders.  For them, as us, the Le diagrams are particularly useful for understanding the dimensions of positroid cells since the dimension becomes the count of the number of plusses.  The ubiquitous Catalan numbers arise for them, as for us in \cite{generalcombinatoricsI}.  The specific subclasses of positroids we work with are, however, different.  This basic methodology of gaining insight from bijections is very common in combinatorics.

@@

Note that the positroid cells associated to the Amplituhedron are of dimension $4|\cP|$. In \cite[section 2]{non-orientable}, Agarwala and Marcott associate a Deodhar component\footnote{Since Deodhar components play no role in this paper, the interested reader is referred to \cite{non-orientable} for more details on this.} to each Wilson loop diagram, and show that the dimension of this Deodhar component is the sum of $|\cP|$ and the dimension of the associated positroid cell.
%\[|\cP| + \text{ the dimension of associated positroid cell}.\]
Combining this with Theorem~\ref{thm dim}, we see that the Deodhar component associated to a Wilson loop diagram is $4|\cP|$-dimensional, showing that the geometry underlying the Wilson loop diagrams is consistent with the geometry underlying the Amplituhedron. It is worth noting that the Amplituhedon is defined by a projection of a set of $4|\cP|$ dimensional positroid cells, while the $4|\cP|$-dimensional cells of the Wilson loop diagram lie in $\mathbb{G}_\R$.
While it is conjectured that the geometry of the Amplituhedron is related to the geometry of the Wilson loop diagrams, in part due to their common physical origin, the $3 |\cP|$ dimensional positroid cells defined by the Wilson loop diagrams do not seem to bear any geometric relationship to the $4|\cP|$ dimensional cells parametrized by the Amplituhedron.
%To  date, there is no evidence to suggest that the $3|\cP|$ dimensional positive cells of $\Gr(|\cP|, n)$ defining the Wilson loop diagrams are related to the $4|\cP|$ dimensional positive cells defining the Amplituhedron.  \sanote{This para adresses point 3 of report Y. Should there be a similar paragraph in the intro? Should it be moved?} \sanote{futzed some  more}

Marcott in \cite[Theorem 8.4]{WLDdim} also gives a proof of the $3|\cP|$-dimensionality of $M(W)$ which is much more geometric. However, it is not easy to track the effect of a particular propagator in that proof. By contrast, our approach is much more constructive, and elucidates how each additional propagator contributes to tthe dimension. By combining Algorithms~\ref{alg:put GN on WLD} and \ref{alg:GN to Le} (converting from Wilson loop diagram to Grassmann necklace, and Grassmann necklace to Le diagram respectively) we explicitly describe the effect of adding another propagator to a Wilson loop diagram in terms of the plusses of the associated Le diagrams. Recall that the dimension of a positroid cell is equal to the number of plusses in its associated Le diagram \cite[Theorem 6.5]{Postnikov}.  Thus understanding the effect of adding another propagator is sufficient to give a recursive proof of the $3|\cP|$-dimensionality of the cells.

>>>>

Wilson loop diagrams, as we will study them, are objects given by a perturbative expansion of the holomorphic Wilson loop in twistor space.
Classically, the interactions of the theory were studied as maximal helicity violating diagrams ($MHV$ diagrams), next-to-maximal helicity violating diagrams and so on ($N^kMHV$ diagrams) in momentum space. Adamo and Mason, in \cite{Adamo:2011pr}, gave an elegant formulation for the $N^kMHV$ diagrams in twistor space. The Wilson loop diagram perspective of this paper is dual to this twistor representation of $MHV$ diagrams.
The objects are closely related and carry the same information.  We choose to emphasize the dual perspective by our focus on the Wilson loop diagrams.

@@

In \cite{wilsonloop}, Agarwala and Marin-Amat uncovered a connection between the Wilson loop diagrams and positroids, defined as matroids that can be realized as an element of $\Gr(k,n)$. Other work has shown a connection between the volume filled by Wilson loop diagrams and the Amplituhedron \cite{casestudy, Amplituhedronsquared}. Specifically, Marcott showed that the space parametrized by the Wilson loop diagrams is  exactly the correct positroid cell in $\Gr(k,n)$, up to a set of measure zero \cite{WLDdim}.

In the above work and elsewhere, it is clear that while the space parametrized by the Wilson loop diagrams is related to the Amplituhedron, they are not the same space, nor is a direct connection between the geometry of the Amplituhedron and the geometry of the Wilson loop diagrams known. For instance, it is conjectured that the Amplituhedron is orientable, whereas Agarwala and Marcott showed in \cite{non-orientable} that the space parametrized by the Wilson loop diagrams as a subspace of $\Gr(k, n+1)$ can be seen as a non-orientable vector bundle over some submanifold of $\Gr(k,n)$. Furthermore, the cells associated to the Wilson loop diagrams are different from the cells associated to BCFW diagrams, which can be seen by comparing the results of \cite{AmplituhedronDecomposition} and \cite{casestudy}. Furthermore, the spaces in the two situations have different dimensions, as we will discuss further below, and examples show \cite{HeslopStewart} that they do not relate by a coordinate projection or boundary map or other straightforward function.  None the less, from their common physical origin, it is believed that once the integrals over these spaces are taken the results will parametrize related spaces \cite{wilsonloop, Amplituhedronsquared}.

@@

This paper continues the program of establishing concrete connections between the combinatorics of the Wilson loop diagram as determined by the physics, and the combinatorics of the associated positroid cell as identified in \cite{wilsonloop}. In this paper, we give an algorithm to read the Grassmann necklace, one of the standard combinatorial characterizations of the positroid cell, directly off the Wilson loop diagram. We also show that the positroid cells associated to Wilson loop diagrams are all $3k$-dimensional, where $k$ is the number of propagators in the diagram; by contrast, the BCFW cells are all $4k$-dimensional \cite{Arkani-Hamed:2013jha}. In \cite{non-orientable} the authors noted that the Wilson loop diagrams only correspond to positroid cells when the gauge vector is ignored. Including the gauge vector defines a $4k$ dimensional subspace of $\mathbb{G}_\R(k,n+1)$.

@@

The matrix $C(W)$ defines a positroid cell associated to $W$ in the positive Grassmannian $\Gr(n, |\cP|)$, which we denote by $\Sigma(W)$ \cite[Theorem 3.41]{wilsonloop}. Marcott \cite[Theorem 8.4]{WLDdim} showed that  the matrix $C(W)$ parametrizes a space whose closure is exactly the closure of $\Sigma(W)$. That is, the space defined by $C(W)$ and the cell $\Sigma(W)$ agree up to a set of measure $0$. Most of the work in this paper and in the previous paper in the series \cite{generalcombinatoricsI} focuses on characterizing the positroid cell associated to each Wilson loop diagram $W$.

@@

Postnikov showed in \cite{Postnikov} that the positroids of $\Gr(k,n)$ are indexed by many different collections of objects, each with their own advantages and disadvantages. The two most suited to our purposes are Grassmann necklaces \cite[section 16]{Postnikov} and Le diagrams \cite[section 6]{Postnikov}, which we discuss. In order to do so, we first need some preliminary definitions.

@@

In \cite{wilsonloop}, Agarwala and Marin-Amat showed that $M(W)$ can also be read directly from the Wilson loop diagram itself.

@@

By relating the behavior of the propagators in $W$ to the matroidal properties of $M(W)$, Agarwala and Marin-Amat also showed that $M(W)$ is a positroid whenever $W$ is admissible \cite[Corollary 3.39]{wilsonloop}.

@@

The flavour of the calculations is similar to\footnote{Thanks to a referee for pointing this paper out.} \cite{CGdyck} as in both our paper and theirs the bijections are very explicit with a similar iterative flavour using Le diagrams and Grassmann necklaces related to another class of objects, for us Wilson loop diagrams and for them unit interval orders.  For them, as us, the Le diagrams are particularly useful for understanding the dimensions of positroid cells since the dimension becomes the count of the number of plusses.  The ubiquitous Catalan numbers arise for them, as for us in \cite{generalcombinatoricsI}.  The specific subclasses of positroids we work with are, however, different.  This basic methodology of gaining insight from bijections is very common in combinatorics.

@@

Note that the positroid cells associated to the Amplituhedron are of dimension $4|\cP|$. In \cite[section 2]{non-orientable}, Agarwala and Marcott associated a Deodhar component\footnote{Since Deodhar components play no role in this paper, the interested reader is referred to \cite{non-orientable} for more details on this.} to each Wilson loop diagram, and showed that the dimension of this Deodhar component is the sum of $|\cP|$ and the dimension of the associated positroid cell.
%\[|\cP| + \text{ the dimension of associated positroid cell}.\]
Combining this with Theorem~\ref{thm dim}, we see that the Deodhar component associated to a Wilson loop diagram is $4|\cP|$-dimensional, showing that the geometry underlying the Wilson loop diagrams is consistent with the geometry underlying the Amplituhedron. It is worth noting that the Amplituhedon is defined by a projection of a set of $4|\cP|$ dimensional positroid cells, while the $4|\cP|$-dimensional cells of the Wilson loop diagram lie in $\mathbb{G}_\R$.
While it is conjectured that the geometry of the Amplituhedron is related to the geometry of the Wilson loop diagrams, in part due to their common physical origin, the $3 |\cP|$ dimensional positroid cells defined by the Wilson loop diagrams do not seem to bear any geometric relationship to the $4|\cP|$ dimensional cells parametrized by the Amplituhedron.
%To  date, there is no evidence to suggest that the $3|\cP|$ dimensional positive cells of $\Gr(|\cP|, n)$ defining the Wilson loop diagrams are related to the $4|\cP|$ dimensional positive cells defining the Amplituhedron.  \sanote{This para adresses point 3 of report Y. Should there be a similar paragraph in the intro? Should it be moved?} \sanote{futzed some  more}

Marcott in \cite[Theorem 8.4]{WLDdim} also gave a proof of the $3|\cP|$-dimensionality of $M(W)$ which is much more geometric. However, it is not easy to track the effect of a particular propagator in that proof. By contrast, our approach is much more constructive, and elucidates how each additional propagator contributes to tthe dimension. By combining Algorithms~\ref{alg:put GN on WLD} and \ref{alg:GN to Le} (converting from Wilson loop diagram to Grassmann necklace, and Grassmann necklace to Le diagram respectively) we explicitly describe the effect of adding another propagator to a Wilson loop diagram in terms of the plusses of the associated Le diagrams. Recall that the dimension of a positroid cell is equal to the number of plusses in its associated Le diagram \cite[Theorem 6.5]{Postnikov}.  Thus understanding the effect of adding another propagator is sufficient to give a recursive proof of the $3|\cP|$-dimensionality of the cells.
